
---
title: "peakProminence Package Vignette"
author: "Author Name"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{peakProminence Package Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Set CRAN mirror to cloud R-project
options(repos = c(CRAN = "https://cloud.r-project.org"))

```

```{r}
library(peakProminence)
```


```{r, include = FALSE}
# Set CRAN mirror to cloud R-project
options(repos = c(CRAN = "https://cloud.r-project.org"))

```

# Introduction
The peakProminence package provides tools for detecting peaks in 1D or 2D data and calculating the prominence of those peaks. Prominence measures how much a peak stands out relative to its surroundings. This package is designed to be useful for analyzing time-series or signal data to identify important features.

In this vignette, we will demonstrate how to:

1. Simulate data

2. Detect peaks in the data

3. Calculate the prominence of detected peaks

4. Visualize the results using built-in plotting functions

# Installation
You can install the peakProminence package directly from GitHub as follows:

```{r}
# Install devtools if you don't have it
install.packages("devtools")

# Install peakProminence from GitHub
devtools::install_github("HebaAleterji/peak-prominence-calculator/peakProminence")

```

# Simulating Data

We'll begin by simulating some data that represents a noisy signal. In this case, we use a sum of Gaussian functions to create peaks at different positions with random noise.

```{r}
# Simulate some data
set.seed(13)
n_comp <- 12
means <- runif(n_comp, -4, 11)
weights <- runif(n_comp, 0.25, 1.1)
weights <- weights / sum(weights)
sigmas <- runif(n_comp, 0.5, 1.5)
x <- seq(-6, 13, length.out = 100)
y <- rep(0, 100)
for (i in 1:n_comp) {
  y <- y + weights[i] * dnorm(x, means[i], sigmas[i])
}

# Plot the simulated data
plot(x, y, type = "l", main = "Simulated Signal", xlab = "Index", ylab = "Value")

```

# Detecting Peaks

To detect peaks in the simulated signal, we use the detect_peaks() function. This function identifies local maxima in the data and returns their positions and heights.

```{r}
# Detect peaks in the simulated data
peaks <- detect_peaks(y)

# Print the detected peaks
print(peaks)

```
# Calculating Prominence
Once the peaks are detected, we can calculate their prominence using the calculate_prominence() function. Prominence measures how much a peak stands out relative to its surrounding valleys.

```{r}
# Calculate prominence for the detected peaks
prominence <- calculate_prominence(peaks, y)

# Print the calculated prominence values
print(prominence)

```

# Combining Peak Information
The get_peak_info() function combines the positions, heights, and prominence values of detected peaks into a single data frame for easier analysis.

```{r}

# Get a data frame with peak information
peak_info <- get_peak_info(peaks, prominence)

# Display the peak information
print(peak_info)

```

# Visualizing Peaks
We can visualize the detected peaks on top of the original signal using the plot_peaks() function.

```{r}

# Plot the original signal and highlight the detected peaks
plot_peaks(peaks, y)

```
# Visualizing Prominence

Similarly, we can visualize the prominence of each detected peak using the plot_prominence() function. This function plots vertical lines indicating how far down each peak's prominence extends.

```{r}
# Plot the prominence of each peak
plot_prominence(peaks, prominence, y)

```

# Conclusion

The peakProminence package offers a simple and effective way to detect peaks and calculate their prominence in both 1D and 2D datasets. In this vignette, we have demonstrated the core functionality of the package using a simulated signal. These tools can be applied to a wide variety of real-world problems where identifying and analyzing important features in time-series or signal data is essential.
